---
title: "On Computing Dynamic SWATH Windows"
author: "Witold E. Wolski, Christian Trachsel, and Christian Panse"
date: "`r doc_date()`"
package: "`r pkg_ver('specL')`"
vignette: >
  %\VignetteIndexEntry{Computing Dynamic SWATH Windows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output: 
  BiocStyle::html_document
---

# Introduction

Function optimizes Extraction windows so we have the same number of precursor per window. To do it uses spectral library or nonredundant blib.

## Requirements

* Mass range can be specified
* Maximal window size can be specified (windows should not be to large because of collision energy optimization (personal communication by Bernd R.))
* Minimal window size can be specified
* target number of windows can be specified.
* boundaries between windows are placed in regions were no precursors are observed.

# Prerequisites

`r Biocpkg("specL")` contains a function `specL::cdsw`.

```{r echo=FALSE}
library(specL)
```


```{r}

.cdsw_compute_breaks <- function(x, n){
  q <- quantile(x, seq(0, 1, length = n + 1))
  q[1] <- q[1] - 0.5
  q[length(q)] <- q[length(q)] + 0.5
  q <- round(q)
}
cdsw <-
function(x, n = 20, overlap = 1.0, ...) {
  if (class(x) == "psmSet") {
    x <- unlist(lapply(x, function(x) {
      x$pepmass
    }))
  } else if (class(x) == 'specLSet') {
    x <- unlist(lapply(x@ionlibrary, function(xx) {
      xx@q1
    }))
  }
  # x should be numeric
  if (class(x) != "numeric") {
    warning("can not compute quantils. 'x' is not numeric.")
    return (NULL)
  }
  
  q <- .cdsw_compute_breaks(x, n)
  
  idx <- 1:n
  from <- q[idx] - overlap * 0.5
  
  to <- q[idx + 1] + overlap * 0.5
  width <- 0.5 * (to - from)
  mid <- from + width
  h <- hist(x,
            breaks = q, ...)
  
  data.frame(from, to, mid, width, counts = h$counts)
}
```

```{r fig.retina=3, warning=FALSE}
cdsw(peptideStd, freq=TRUE, overlap = 0, main = "peptideStd", xlab='pepmass')
```



# Session info

Here is the output of `sessionInfo()` on the system on which this
document was compiled:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

