---
title: "On Computing Dynamic SWATH Windows"
author: "Witold E. Wolski, Bernd Roschitzki, Christian Trachsel, and Christian Panse"
date: "`r doc_date()`"
package: "`r pkg_ver('specL')`"
vignette: >
  %\VignetteIndexEntry{Computing Dynamic SWATH Windows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output: 
  BiocStyle::html_document
---

__!!!Caution work in progress!!!__

# Introduction


Function optimizes Extraction windows so we have the same number of precursor per window. To do it uses spectral library or nonredundant blib.

## Requirements

* Mass range can be specified (*mass\_range*)
* Maximal window size can be specified (*max\_window\_size*). This is because windows should not be to large because of optimal collision energy (personal communication by Bernd R.).
* Minimal window size can be specified (*min\_window\_size*).
* target number of windows can be specified (*nr\_windows*).
* boundaries between windows are placed in regions were no precursors are observed.

# Prerequisites

`r Biocpkg("specL")` contains a function `specL::cdsw`.

```{r echo=FALSE}
library(specL)
```


```{r}

.cdsw_compute_breaks <- function(x, n){
  q <- quantile(x, seq(0, 1, length = n + 1))
  q[1] <- q[1] - 0.5
  q[length(q)] <- q[length(q)] + 0.5
  q <- round(q)
}
cdsw <-
function(x, n = 20, overlap = 1.0, ...) {
  if (class(x) == "psmSet") {
    x <- unlist(lapply(x, function(x) {
      x$pepmass
    }))
  } else if (class(x) == 'specLSet') {
    x <- unlist(lapply(x@ionlibrary, function(xx) {
      xx@q1
    }))
  }
  # x should be numeric
  if (class(x) != "numeric") {
    warning("can not compute quantils. 'x' is not numeric.")
    return (NULL)
  }
  
  q <- .cdsw_compute_breaks(x, n)
  
  idx <- 1:n
  from <- q[idx] - overlap * 0.5
  
  to <- q[idx + 1] + overlap * 0.5
  width <- 0.5 * (to - from)
  mid <- from + width
  h <- hist(x,
            breaks = q, ...)
  
  data.frame(from, to, mid, width, counts = h$counts)
}
```

```{r fig.retina=3, warning=FALSE}
cdsw(peptideStd, freq=TRUE, overlap = 0, main = "peptideStd", xlab='pepmass')
```


# optimze windows

```{r}

objective <- function(splits, data){
  counts <- hist(data, breaks=splits,plot=FALSE)$counts
  nbins<-length(splits)-1
  optimumN <- length(data)/(length(splits)-1)
  optimumN<-rep(optimumN,nbins)
  
  score2 <-sqrt(sum((counts - optimumN)^2))
  score1 <- sum(abs(counts - round(optimumN)))
  return(list(score1=score1,score2 = score2, counts=counts, optimumN=round(optimumN)))
}

hardconstrain <- function(splits, minwindow = 5, maxwindow=50){

  difsp<-diff(splits)
  return(sum(difsp >= minwindow) == length(difsp) & sum(difsp <= maxwindow) == length(difsp))
}

generatesplits <- function(start,end, nrbins = 10){
  sample<-runif(nrbins-1,min=start, max=end)
  c(start,sort(round(sample)),end)
}

main <- function(data, nrbins=50, minwindow=5, maxwindow=50){
  start <- floor(min(data))
  end <- ceiling(max(data))
  start
  end
  set.seed(1)
  res<-lapply(1:100, function(x){generatesplits(start, end,nrbins= nrbins)} )
  sum(sapply(res, function(x){hardconstrain(x, minwindow=0, maxwindow=50)}))

}

```

# Distribution Mixing

```{r}

xx <- sapply(peptideStd, function(x){x$pepmass})


nrwindows = 30

hist(xx)
nbins=35

range(xx)
xx<-xx[xx >=310 & xx<1250]
hist(xx)
qqs <- quantile(xx,probs = seq(0,1,by=1/(nbins-1)))
plot(1:nbins,qqs,type="b")
unif <- seq(min(xx),max(xx),length=(nbins))
lines(1:nbins,unif,col=2,type="b")

hardconstrain(unif,minwindow = 5, maxwindow = 80)
hardconstrain(qqs,minwindow = 5, maxwindow = 80)

bb <- xx

while(!hardconstrain(qqs,minwindow = 5, maxwindow = 80)){
  xx2<-runif(10, min=min(xx), max=max(xx))
  bb<-c(bb,xx2)
  #hist(bb)
  qqs<-quantile(bb,probs = seq(0,1,by=1/(nbins-1)))
  lines(1:nbins,qqs,type="l",col=3)
}

lines(1:nbins,qqs,type="b",col=4)
diff(qqs)


```


# Session info

Here is the output of `sessionInfo()` on the system on which this
document was compiled:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```





