---
title: "On Computing Dynamic SWATH Windows"
author: "Witold E. Wolski, Bernd Roschitzki, Christian Trachsel, and Christian Panse"
date: "`r doc_date()`"
package: "`r pkg_ver('specL')`"
vignette: >
  %\VignetteIndexEntry{Computing Dynamic SWATH Windows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output: 
  BiocStyle::html_document
---

__!!!Caution work in progress!!!__

# Introduction


Function optimizes Extraction windows so we have the same number of precursor per window. To do it uses spectral library or nonredundant blib.


# Prerequisites

`r Biocpkg("specL")` contains a function `specL::cdsw`.

```{r echo=FALSE}
library(specL)

exampledata <- ms1.p2069

```

# Classical Method based on `quantile` 
```{r}

.cdsw_compute_breaks <- 
  function(xx, nbins){
  q <- quantile(xx, seq(0, 1, length = nbins + 1))
  q[1] <- q[1] - 0.5
  q[length(q)] <- q[length(q)] + 0.5
  q <- round(q)
}

cdsw <-
  function(x, n = 20, overlap = 1.0, FUN, ...) {
  if (class(x) == "psmSet") {
    x <- unlist(lapply(x, function(x) {
      x$pepmass
    }))
  } else if (class(x) == 'specLSet') {
    x <- unlist(lapply(x@ionlibrary, function(xx) {
      xx@q1
    }))
  }
  # x should be numeric
  if (class(x) != "numeric") {
    warning("can not compute quantils. 'x' is not numeric.")
    return (NULL)
  }
  
  q <- FUN(xx=x, nbins=n)
  
  idx <- 1:n
  from <- q[idx] - overlap * 0.5
  
  to <- q[idx + 1] + overlap * 0.5
  width <- 0.5 * (to - from)
  mid <- from + width
  h <- hist(x, breaks = q, ...)
  
  data.frame(from, to, mid, width, counts = h$counts)
}
```

```{r fig.retina=3, warning=FALSE}


cdsw(exampledata, 
     freq=TRUE, 
     overlap = 0, 
     main = "peptideStd", xlab='pepmass', FUN=.cdsw_compute_breaks)
```

# Iterative Distribution Mixing based cdsw

## Requirements

- Mass range can be specified (*mass\_range*)
- Maximal window size can be specified (*max\_window\_size*). This is because windows should not be to large because of optimal collision energy (personal communication by Bernd R.).
- Minimal window size can be specified (*min\_window\_size*).
- target number of windows can be specified (*nr\_windows*).
- boundaries between windows are placed in regions were no precursors are observed.


## Contrains and Objective Function

```{r}

.cdsw_objective <- function(splits, data){
  counts <- hist(data, breaks=splits,plot=FALSE)$counts
  nbins<-length(splits)-1
  optimumN <- length(data)/(length(splits)-1)
  optimumN<-rep(optimumN,nbins)
  
  score2 <-sqrt(sum((counts - optimumN)^2))
  score1 <- sum(abs(counts - round(optimumN)))
  return(list(score1=score1,score2 = score2, counts=counts, optimumN=round(optimumN)))
}

.cdsw_hardconstrain <- function(splits, minwindow = 5, maxwindow=50){

  difsp<-diff(splits)
  return(sum(difsp >= minwindow) == length(difsp) & sum(difsp <= maxwindow) == length(difsp))
}

```

## Construction Heuristic

```{r}
.cdsw_compute_sampling_breaks <- function(xx, nbins=35, maxwindow=150, plot=TRUE){
  nbins<-nbins+1
  it_count <- 0
  error <- 0
  #xx<-xx[xx >=310 & xx<1250]
  # TODO(wew): there is something insconsitent with the nbins parameter
  qqs <- quantile(xx,probs = seq(0,1,by=1/(nbins-1)))
  
  plot(1:nbins, qqs, type="b" , 
       sub=".cdsw_compute_sampling_breaks")
  legend("topleft", legend = c(paste("maxwindow = ", maxwindow),
        paste("nbins = ", nbins) ))
  unif <- seq(min(xx),max(xx),length=(nbins))
  
  lines(1:nbins,unif,col=2,type="b")
  
  .cdsw_hardconstrain(unif,minwindow = 5, maxwindow)
  .cdsw_hardconstrain(qqs,minwindow = 5, maxwindow)
  
  bb <- xx
  
  while(!.cdsw_hardconstrain(qqs,minwindow = 5, maxwindow)){
    it_count <- it_count + 1
    xx2<-runif(round(length(xx)/10), min=min(xx), max=max(xx))
    bb<-c(bb,xx2)
   
    qqs<-quantile(bb,probs = seq(0,1,by=1/(nbins-1)))
    lines(1:nbins,qqs,type="l", col="#00DD00AA")
    
    error[it_count] <-.cdsw_objective(qqs, xx)$score1
    
  }
  
  lines(1:nbins,qqs,type="b", col="#FF1111AA")
  
  plot(error, xlab="number of iteration", sub=".cdsw_compute_sampling_breaks")
  #list(qqs=qqs, error=error)

  #qqs[length(qqs)] <- round(max(xx)+10)
  
  qqs <- as.numeric(sort(round(qqs)))
  qqs[1]  <- qqs[1] - 0.5
  
  qqs[length(qqs)]  <- qqs[length(qqs)] + 0.5
 
  round(qqs, 1)
}
```


## Evaluation
### Comparizon to Classical Approach
```{r fig.height=8, fig.width=8, fig.retina=3}
op <-par(mfrow=c(2,2))


cdsw(exampledata, 
     freq=TRUE,
     plot=TRUE,
     overlap = 0, 
     sub='sampling based',
     main = "peptideStd", xlab='pepmass', FUN=function(...){.cdsw_compute_sampling_breaks(...,maxwindow = 100)})


cdsw(exampledata, 
     freq=TRUE,
     plot=TRUE,
     overlap = 0, 
      sub='quantile based',
     main = "peptideStd", xlab='pepmass', FUN=.cdsw_compute_breaks)

```

### Comparizon using different MaxWindowSize
```{r fig.height=10, fig.width=8, fig.retina=3, warning=FALSE}
op <-par(mfrow=c(4,3))
res <- lapply(c(75,150,300, 800), function(mws){
  cdsw(exampledata, 
     freq=TRUE,
     plot=TRUE,
     overlap = 0, 
     main = paste("max window size", mws), xlab='pepmass', 
     FUN=function(...){
       .cdsw_compute_sampling_breaks(...,maxwindow = mws)
       })
})
```

# Chaning nr bins

```{r fig.height=10, fig.width=8, fig.retina=3, warning=FALSE}
op <-par(mfrow=c(4,3))
res <- lapply(c(20,25,30, 40), function(nbins){
  cdsw(exampledata, 
     freq=TRUE,
     plot=TRUE,
     n=nbins,
     overlap = 0, 
     main = paste("nr bins", nbins), xlab='pepmass', 
     FUN=function(...){
       .cdsw_compute_sampling_breaks(...,maxwindow = 100)
       })
})
```

# Session info

Here is the output of `sessionInfo()` on the system on which this
document was compiled:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```





